\begin{abstract}
We present TransitDuet, a bi-level reinforcement learning framework that jointly optimises adaptive bus target-headway planning and real-time holding control.
This joint problem is fundamentally harder than either task in isolation: the planning and control layers operate on different timescales, observe different state spaces, and interact through delayed, asynchronous feedback---properties that violate the synchronous co-execution assumption underlying existing hierarchical reinforcement learning methods.
TransitDuet addresses this through three mechanisms.
First, a goal-conditioned cross-level coupling in the spirit of HIRO --- adopted, not re-derived, and selected via end-to-end comparison against a literal cross-advantage augmentation and a three-channel launch-time-perturbation alternative (Section V): the upper outputs a per-dispatch target-headway shift, and the lower's Lagrangian cost is computed against the resulting goal headway, with no upper-advantage injection into the lower's transitions. This avoids the covariate shift that destabilises a literal cross-advantage augmentation while keeping the upper--lower interaction unidirectional and stable.
Second, a measurement-driven online gradient descent ($\thetavec$-OGD) on the upper-reward weights adaptively penalises fleet overshoot --- this is a soft adaptive penalty rather than a strict policy projection, but it removes the need for a hand-tuned overshoot coefficient and converges to a regime where the constraint is rarely violated in expectation.
Third, Lagrangian relaxation within the lower policy penalizes headway deviations, converting a hard service-regularity requirement into a soft dual objective.
We evaluate TransitDuet in a calibrated microsimulation of a high-frequency bus corridor with stochastic demand and traffic.
A complementary technique, target-policy-corrected lower SAC (TPC), uses importance-reweighting to align the lower's training distribution with its eventual deployment distribution; combined with the goal-conditioned coupling and trained for 300 episodes per seed (search baselines: 320 episodes), TransitDuet attains the best numerical mean of any method on all four reported metrics (wait, CV, overshoot, composite) under a near-unified, approximately budget-matched evaluation protocol; the overshoot margin over CMA-ES is $0.03$ and we treat it as within seed noise rather than as a substantive advantage. Across three random seeds and an elastic fleet budget sampled per episode from $[8, 16]$, TransitDuet reduces composite cost by $8.2\%$ ($1.418$ vs $1.544$) relative to the strongest hand-tuned fixed timetable, $4.7\%$ relative to the strongest search baseline (CMA-ES at $1.488$), and is markedly more reliable than GA, which has a 1-in-3 cold-start failure mode that pushes its mean composite to $1.800$. Seed variance on composite cost is the lowest of any method (std $0.016$, vs $0.022$ for CMA-ES, $0.040$ for Fixed, and $0.458$ for GA), and a single trained policy per seed produces an operational Pareto frontier across the nine fleet sizes $\Nfleet \in [8, 16]$ (composite-knee at $\Nfleet{=}14$, no retraining required). Generalisation degrades gracefully: wait rises by $16\%$ at $\sigma_{\mathrm{route}}{=}2.0$ and $27\%$ at $\sigma_{\mathrm{route}}{=}3.0$ (double the training stochasticity). We acknowledge the small seed count ($n{=}3$) limits significance testing; we report ranges and standard deviations throughout and avoid pairwise significance claims.
\end{abstract}
