\section{Conclusion}
\label{sec:conclusion}

This paper introduced TransitDuet, a bi-level reinforcement learning framework that unifies adaptive bus target-headway planning and real-time holding control within a single trainable system.
The upper policy learns a per-dispatch target-headway shift $\delta_t \in [-120, 120]$\,s on top of a fixed nominal timetable, while the lower policy regulates individual holding actions at station arrivals---two tasks that have traditionally been addressed in isolation despite their tight operational coupling.

The technical core is a goal-conditioned cross-level coupling (Section~\ref{sec:tap}) in the spirit of HIRO~\citep{nachum2018data}: the upper outputs a per-dispatch target-headway shift, and the lower's Lagrangian cost is evaluated against the resulting goal headway, with no upper-advantage injection into the lower's transitions.
We compared this end-to-end against a literal cross-advantage augmentation following HAAR~\citep{li2019hierarchical} (with a PIPER-style reachability gate~\citep{singh2024piper}) and against a three-channel coupling that perturbs the launch time and supplies a hindsight reward credit; the goal-conditioned variant attained the best mean composite cost and the lowest seed variance among the three.
A target-policy-corrected lower SAC (TPC, Section~\ref{sec:tpc}) provides a complementary training-time correction that prevents residual covariate shift in the lower's training distribution.
Operational constraints are handled through complementary mechanisms: online gradient descent projection ($\thetavec$-OGD) reweights the upper-level fleet-overshoot penalty, while Lagrangian relaxation maintains headway regularity at the lower level.

Experiments on a stochastic bus corridor microsimulation with 22 stations and 264 daily trips show that, under a near-unified protocol (TransitDuet trains for 300 episodes per seed; search baselines for 320 = 20 warmup + 300 search-eval, with identical lower hyperparameters), TransitDuet matches or improves on every baseline on its strongest metrics: composite cost $1.418$ vs $1.488$ for the strongest search baseline (CMA-ES, $-4.7\%$) and $1.544$ for the strongest hand-tuned fixed timetable ($-8.2\%$), with the best numerical mean on all four reported metrics (the overshoot margin over CMA-ES is within seed noise and we do not claim a substantive overshoot advantage). GA in this protocol shows a 1-in-3 cold-start failure mode that pushes its composite to $1.800$, illustrating the unreliability of population-based search without a pre-trained lower at the same budget. From a single training run the learned policy produces a Pareto frontier across nine fleet budgets ($\Nfleet \in [8, 16]$), removing the need for separate optimization per budget.
Architectural and training ablations identify the goal-conditioned coupling (Section~\ref{sec:tap}) as the dominant design choice within the bi-level framework, while elastic-fleet sampling, holding-feedback, and the trip-lifecycle hindsight credit are the three auxiliary components whose removal materially increases composite cost ($\Delta \geq 9.7\%$); TPC's effect is mostly to tighten seed variance. The remaining components address out-of-distribution properties measured separately in the generalization study.
Without retraining, the learned policy retains its performance under route stochasticity for $\sigma \in \{0.5, 1.0\}$ (wait is comparable to or slightly below training-$\sigma$), degrades by $16\%$ at $\sigma = 2.0$, and remains operational at $\sigma = 3.0$ with a $27\%$ wait increase relative to the training distribution.

Several directions merit future investigation.
First, extending TransitDuet to multi-route networks with shared fleets and transfer coordination would test whether the hierarchical structure scales to the combinatorial action spaces of realistic transit systems.
Second, real-world deployment---potentially through a calibrated digital twin with domain randomization---would validate the framework's robustness to phenomena not captured by the current simulator, including traffic signal interactions and driver behavior heterogeneity.
Third, integrating real-time demand forecasting and modeling passenger demand elasticity would close the loop between service quality and ridership.
Finally, coupling TransitDuet with passenger information systems---broadcasting predicted arrival times based on the learned holding policy---could amplify the welfare benefits by reducing perceived waiting time alongside actual waiting time.
