\section{Hyperparameter details}
\label{app:hyperparams}

\Cref{tab:hyperparams} lists all hyperparameters used in the experiments. Values correspond to the default configuration; no per-experiment tuning was performed.

\begin{table*}[!t]
\centering
\caption{Complete hyperparameter specification.}
\label{tab:hyperparams_full}
\small
\begin{tabular}{@{}llll@{}}
\toprule
\textbf{Group} & \textbf{Parameter} & \textbf{Value} & \textbf{Description} \\
\midrule
\multirow{4}{*}{Environment}
  & \texttt{route\_sigma}        & 1.5    & Log-normal speed noise $\sigma$ \\
  & \texttt{max\_agent\_num}     & 25     & Hard cap on concurrent buses (env-side, $> \Nfleet^{\max}$) \\
  & \texttt{effective\_trip\_num} & 264   & Total trips per service day \\
  & \texttt{demand\_noise}        & 0.15  & Per-hour demand multiplier std \\
\midrule
\multirow{13}{*}{Upper policy}
  & \texttt{state\_dim}    & 11             & system summary (5) + HoldFB (3) + trip ctx (3); see Eq.~\ref{eq:upper_state} \\
  & \texttt{action\_dim}   & 1              & per-dispatch headway shift $\delta_t \in [-120, 120]$\,s \\
  & \texttt{action\_low}   & $-120$         & Min offset (s) \\
  & \texttt{action\_high}  & $+120$         & Max offset (s) \\
  & \texttt{hidden\_dim}   & 64             & MLP hidden layer width \\
  & \texttt{lr}            & $3 \times 10^{-4}$ & Policy / critic learning rate (Adam) \\
  & \texttt{batch\_size}   & 64             & Replay buffer mini-batch \\
  & \texttt{updates\_per\_episode} & 10     & SAC gradient steps per training episode \\
  & \texttt{replay\_capacity}      & 50{,}000 & Upper transitions stored \\
  & \texttt{ensemble\_size} $K$    & 10     & RE-SAC critic ensemble size \\
  & $\beta_{\mathrm{LCB}}$         & $-2.0$ & LCB coefficient in policy improvement \\
  & $\gamma_{\mathrm U}$           & 0.95   & Discount factor (per-dispatch) \\
  & $\alpha_{\max}^{\mathrm U}$    & 0.05   & Maximum entropy temperature (auto-tuned) \\
  & $\Nfleet^{\min},\Nfleet^{\max}$ & $8, 16$ & Elastic fleet sampling range \\
  & $\Nfleet$ (default)    & 12             & Soft fleet target (overshoot$^2$ penalty) \\
\midrule
\multirow{14}{*}{Lower policy}
  & \texttt{hidden\_dim}     & 64      & MLP hidden layer width \\
  & \texttt{action\_range}   & 60      & Holding time bound (action $\in [-60, 60]$\,s) \\
  & $\climit$                & 0.5     & Lagrangian cost budget \\
  & \texttt{batch\_size}     & 512     & Replay buffer mini-batch \\
  & \texttt{updates\_per\_episode} & 30 & SAC gradient steps per training episode \\
  & \texttt{lr}              & $3 \times 10^{-4}$ & Critic + policy learning rate \\
  & $\lambda_{\text{lr}}$    & $10^{-4}$ & Lagrangian multiplier LR \\
  & $\gamma$                 & 0.99    & Discount factor \\
  & $\tau$                   & 0.005   & Soft target update rate \\
  & \texttt{auto\_entropy}   & true    & Automatic entropy tuning \\
  & $\alpha_{\max}^{\mathrm L}$ & 0.1   & Maximum entropy temperature \\
  & \texttt{ensemble\_size} $K$ & 10   & RE-SAC critic ensemble size \\
  & $\beta_{\mathrm{ood}}$   & 0.1     & Ensemble-spread regulariser weight \\
  & \texttt{reward\_scale}   & 1.0     & Reward scaling factor \\
\midrule
\multirow{6}{*}{Coupling}
  & $e_{\mathrm{warm}}$       & 30 episodes   & Lower-only warmup ($\delta_t = 0$) \\
  & $\alpha_{\mathrm H}$      & 0.5           & Hindsight-credit weight (Eq.~\ref{eq:hindsight_reward}) \\
  & TPC $\epsilon$            & 0.25          & Behaviour-mixture exploratory probability \\
  & TPC $\sigma_{\mathrm{tgt}}$ & 20 s        & Target-mode Gaussian std \\
  & TPC $w_{\max}$            & 5             & IS-weight clip \\
  & TPC EMA $\tau$            & 0.005         & Polyak rate of EMA target upper \\
  & $\eta_{\thetavec}$        & 0.01          & ApproPO $\thetavec$-OGD step size \\
\midrule
\multirow{3}{*}{Training}
  & \texttt{total\_episodes}       & 300     & Total training episodes per seed \\
  & \texttt{save\_freq}            & 50      & Checkpoint frequency (episodes) \\
  & \texttt{replay\_buffer\_size}  & $5 \times 10^{5}$ & Lower replay buffer capacity \\
\bottomrule
\end{tabular}
\end{table*}

%% ---------------------------------------------------------------
\section{Environment details}
\label{app:env}

The simulation models a single bidirectional bus route calibrated from operational data of an urban bus corridor.

\paragraph{Route topology.}
The route comprises $M$ stations connected by inter-station segments.
Each direction (upstream and downstream) traverses all $M$ stations in sequence, with terminal turnaround at each end.
Station positions and inter-station distances are read from operational records (\texttt{stop\_news.xlsx}), and segment-level speed profiles are specified per hour of day (\texttt{route\_news.xlsx}).

\paragraph{Stochastic travel speed.}
Inter-station travel times are driven by a stochastic speed model.
At each route-state update, the speed limit on segment $(i, i{+}1)$ is drawn as
\begin{equation}
  v_{i} = \min\!\bigl(V_{\max},\;\max\bigl(0,\;\lfloor\ln\mathcal{LN}(\mu_i(t),\,\sigma)\rfloor\bigr)\bigr),
\end{equation}
where $\mu_i(t)$ is the historical mean log-speed for segment~$i$ at hour~$t$, $\sigma = 1.5$ is the log-normal dispersion parameter (\texttt{route\_sigma}), and $V_{\max}$ is the segment maximum speed.
This produces realistic speed variability that induces headway deviations even under a well-designed timetable.

\paragraph{Passenger generation.}
Passenger arrivals follow a Poisson process governed by time-varying origin--destination (OD) matrices (\texttt{passenger\_OD.xlsx}).
The OD data specifies hourly demand rates between all station pairs.
At each simulation second, for every origin station with non-zero demand to a given destination, a Poisson draw with rate $\lambda_{od}/3600$ determines the number of new passengers.
Each passenger records an appearance time, an origin, and a destination; waiting time is measured from appearance to boarding.

\paragraph{Bus operations.}
Each bus maintains a state machine with four phases:
\begin{enumerate}
  \item \textbf{Holding} --- the bus is held at a station for the duration specified by the lower-level policy (0--60\,s).
  \item \textbf{Dwelling} --- passengers alight and board. Boarding proceeds until the bus reaches capacity ($C = 50$ passengers) or no more waiting passengers have a valid destination downstream.
  \item \textbf{Travel} --- the bus moves toward the next station at the current stochastic speed. Acceleration ($3\;\text{m/s}^2$) and deceleration ($5\;\text{m/s}^2$) are modeled at departure and arrival.
  \item \textbf{Waiting for action} --- upon arriving at a station, the bus waits for the next holding action from the lower-level policy.
\end{enumerate}
Forward and backward headways are computed with respect to the preceding and following buses in the same direction, and the headway deviation from the target set by the upper policy defines the per-step cost (\Cref{eq:lower_cost}).

\paragraph{Timetable structure.}
A service day consists of 264 trips alternating between the two directions, truncated from a full operational schedule to keep simulation tractable.
At most 25 buses may be concurrently active (\texttt{max\_agent\_num}).
The upper policy outputs a per-dispatch target-headway shift $\delta_t$; the lower's constraint is then computed against the resulting goal headway $h^{\mathrm{target}} + \delta_t$. The nominal launch times of the underlying timetable remain unchanged (no launch-time perturbation in the main HIRO mode).

\paragraph{Episode termination.}
An episode ends when all 264 trips have been launched and every bus has completed its trip (i.e., no bus remains on-route). At this point the measurement vector $\zvec$ (\Cref{eq:measurement}) is computed for the ApproPO $\thetavec$-OGD update.

%% ---------------------------------------------------------------
\section{Network architectures}
\label{app:networks}

All networks use ReLU activations and are trained with Adam. Weight initialization follows PyTorch defaults unless otherwise noted.

\paragraph{Upper policy $\piU$ (Gaussian, scalar action).}
A two-layer MLP maps the 11-dimensional upper state (5 system summary dims + 3 holding-feedback dims + 3 trip-context dims; \cref{eq:upper_state}) to a scalar per-dispatch target-headway shift $\delta_t \in [-120, +120]$\,s:
\begin{center}
\small
\begin{tabular}{lll}
\toprule
\textbf{Layer} & \textbf{Shape} & \textbf{Activation} \\
\midrule
Linear & $11 \to 64$ & ReLU \\
Linear & $64 \to 64$ & ReLU \\
Mean head & $64 \to 1$ & Sigmoid $\to [-120, 120]$\,s \\
Log-std head & $64 \to 1$ & Clamp$[-5, 0]$ \\
\bottomrule
\end{tabular}
\end{center}
The mean and log-std heads are initialized with zero bias to start near $\delta_t = 0$ (no shift relative to the nominal timetable).
During rollout, the pre-squash sample $z \sim \mathcal{N}(\mu, \sigma^2)$ is passed through the sigmoid + affine rescale to give $\delta_t \in [-120, 120]$\,s, which the runner uses as a goal-headway shift (HIRO coupling).
The reparameterization trick is used for policy-gradient evaluation; log-probabilities are computed for the 1-dimensional action.

\paragraph{Lower policy $\piL$ (Squashed Gaussian).}
A two-layer MLP with tanh squashing maps the lower state to a scalar holding time:
\begin{center}
\small
\begin{tabular}{lll}
\toprule
\textbf{Layer} & \textbf{Shape} & \textbf{Activation} \\
\midrule
Linear & $(8{+}K) \to 64$ & ReLU \\
Linear & $64 \to 64$ & ReLU \\
Mean head & $64 \to 1$ & --- \\
Log-std head & $64 \to 1$ & Clamp$[-20, 2]$ \\
\bottomrule
\end{tabular}
\end{center}
The raw sample $z \sim \mathcal{N}(\mu, \sigma^2)$ is squashed via $a = \tanh(z) \times 60$ to produce a holding action in $[-60, 60]$\,s. Negative values are interpreted as a request for early dispatch (clipped at the bus's earliest feasible time); the deployment-time effective range is $[0, 60]$\,s.
Log-probabilities include the standard tanh Jacobian correction \citep{haarnoja2018soft}.
Weights in the output heads are initialized uniformly in $[-3{\times}10^{-3}, 3{\times}10^{-3}]$.
The same network is shared across all buses (parameter sharing), with bus-specific behavior arising from heterogeneous input states.

\paragraph{Reward Q-ensemble $\{Q_k\}_{k=1}^{K}$ (RE-SAC, $K{=}10$).}
The lower replaces the standard twin-Q SAC critic by an ensemble of $K{=}10$ independently-initialised Q-networks; the policy is improved against a lower-confidence-bound (LCB) value $\bar{Q} + \beta_{\mathrm{LCB}} \sigma_Q$ with $\beta_{\mathrm{LCB}} = -2$, computed across the ensemble at every gradient step. Each ensemble member shares the same MLP shape:
\begin{center}
\small
\begin{tabular}{lll}
\toprule
\textbf{Layer} & \textbf{Shape} & \textbf{Activation} \\
\midrule
Linear & $(8{+}K{+}1) \to 64$ & ReLU \\
Linear & $64 \to 64$ & ReLU \\
Linear & $64 \to 1$ & --- \\
\bottomrule
\end{tabular}
\end{center}
Each $Q_k$ takes the concatenation of the lower state and the scalar action as input. All members share a common mean target $(1/K)\sum_k \bar{Q}_k$ during temporal-difference learning to prevent ensemble divergence; per-Q MSE is augmented with a small ensemble-spread penalty $\beta_{\mathrm{ood}}\,\mathrm{std}_k(Q_k)$ ($\beta_{\mathrm{ood}}{=}0.1$) that encourages disagreement on out-of-distribution actions. All ensemble members are accompanied by exponential moving average target networks with Polyak rate $\tau = 0.005$. (We use a Q-ensemble rather than a twin Q to obtain calibrated lower-confidence bounds; the twin-Q variant in standard SAC corresponds to $K{=}2$ without the LCB.)

\paragraph{Cost Q-network $Q_c$.}
A single MLP with the same MLP shape as one ensemble member estimates the expected cumulative cost (headway deviation) under the current policy:
\begin{center}
\small
\begin{tabular}{lll}
\toprule
\textbf{Layer} & \textbf{Shape} & \textbf{Activation} \\
\midrule
Linear & $(8{+}K{+}1) \to 64$ & ReLU \\
Linear & $64 \to 64$ & ReLU \\
Linear & $64 \to 1$ & --- \\
\bottomrule
\end{tabular}
\end{center}
The cost critic is trained with the same target-network mechanism and soft update rate as the reward critics.
Its output enters the policy loss through the Lagrangian multiplier $\lambda$: the policy minimizes $\alpha \log \pi - Q + \lambda Q_c$.

%% ---------------------------------------------------------------
\section{Auxiliary hindsight credit: derivation and contrast with cross-advantage}
\label{app:coupling_derivation}

The main cross-level coupling in TransitDuet is the goal channel of \cref{sec:tap}: the upper outputs a per-dispatch target-headway shift that resets the lower's Lagrangian target. This appendix derives the auxiliary hindsight credit on the upper reward (Eq.~\ref{eq:hindsight_reward} in \cref{sec:tap}), an upper-only signal that supports per-dispatch credit assignment, and contrasts it with a literal asynchronous extension of the cross-advantage objective of \citet{pan2024roboduet}. The literal cross-advantage extension corresponds to one of the architectural alternatives we evaluated end-to-end (Section~\ref{sec:main_results}); the goal-conditioned design we ultimately adopt sidesteps the issues described below by avoiding upper-advantage injection into the lower altogether.

\subsection{Synchronous cross-advantage (RoboDuet, background)}

In the synchronous setting of \citet{pan2024roboduet}, two agents act at every time step $t$ on a shared state $s_t$. The cross-advantage augmented objective is
\begin{equation}
  \mathcal{L}_{\text{upper}}^{\text{sync}} = -\mathbb{E}_{t}\bigl[(A_{\mathrm{U}}(s_t, a_t^{\mathrm{U}}) + \beta\, A_{\mathrm{L}}(s_t, a_t^{\mathrm{L}}))\,\nabla\!\log\piU(a_t^{\mathrm{U}} \mid s_t)\bigr],
  \label{eq:roboduet_upper}
\end{equation}
which is well-defined because both advantages are evaluated at the same shared state-action pair.

\subsection{The asynchrony challenge}

In TransitDuet, $\piU$ acts at dispatch events $\{t_1^{\mathrm{U}}, \ldots, t_N^{\mathrm{U}}\}$ while $\piL$ acts at station arrivals $\{t_1^{\mathrm{L}}, \ldots, t_M^{\mathrm{L}}\}$ with $M \gg N$, on disjoint event sets and different state spaces. There is no single time step at which $\AU$ and $\AL$ are jointly defined on a shared state, so the direct sum in Eq.~\eqref{eq:roboduet_upper} is undefined.

\subsection{Trip-lifecycle alignment and the realised gap}

We resolve this through a \emph{trip-lifecycle} alignment. Each upper action $a_{\mathrm{U},i}$ at dispatch $t_i^{\mathrm{U}}$ initiates a trip $\tau_i$ launched at the baseline time $t_{\mathrm{base},i}$ with a dispatch-specific target headway $\htarget + \delta_{t,i}$. The realised inter-dispatch gap is the actual elapsed time between consecutive bus departures as carried out by the lower controller (which may hold individual buses to track its target):
\begin{equation}
  g_i = t_{\mathrm{actual},i+1} - t_{\mathrm{actual},i},
\end{equation}
computed per direction so that buses operating on opposite legs do not interleave. The per-direction mean and standard deviation of $\{g_i\}$ define the deviation
\begin{equation}
  \Delta_i = \frac{|g_i - \bar{g}|}{\sigma_g}.
  \label{eq:gap_dev}
\end{equation}

\subsection{Hindsight-credited upper reward}

The upper reward at dispatch $i$ is the sum of the episode-level system reward (a belief-weighted MORL scalarisation of wait, fleet overshoot, and headway CV; Eq.~\ref{eq:morl_reward}) and a per-dispatch hindsight credit derived from $\Delta_i$:
\begin{equation}
  R_{\mathrm{U}}^{(i)} = R_{\mathrm{sys}}(\zvec) - \alpha_{\mathrm{H}} \cdot \frac{\Delta_i - \bar{\Delta}}{\sigma_{\Delta}}, \qquad \alpha_{\mathrm{H}} = 0.5.
  \label{eq:hindsight_reward_app}
\end{equation}
The credit is recentred per episode so that dispatches with above-average gap deviation receive a negative credit and dispatches with below-average deviation receive a positive one; the system reward is shared across all dispatches in an episode.

\subsection{Why we did not use literal cross-advantage}

A literal asynchronous extension of Eq.~\eqref{eq:roboduet_upper} would aggregate per-trip lower advantages into the upper objective:
\begin{equation}
  \tilde{R}^{\text{lit},(i)}_{\mathrm{U}} = R_{\mathrm{U}}^{(i)} + \beta \cdot \tfrac{1}{|\mathcal{J}(i)|}\sum_{j \in \mathcal{J}(i)} A_{\mathrm{L}}(s_{\mathrm{L},j}, a_{\mathrm{L},j}),
  \label{eq:cross_adv_literal}
\end{equation}
where $\mathcal{J}(i)$ is the set of lower-level transitions during trip~$i$. We tried this in early prototypes and found it unstable. Two diagnostics: (i) the per-step lower advantage and the per-dispatch upper advantage differ in scale by 1--2 orders of magnitude across a trip lifecycle, so a single scalar $\beta$ either suppresses cross-level signal or destabilises lower training; (ii) the lower's advantage estimator uses an LCB-discounted ensemble Q (\cref{sec:lower}), which is a non-linear function of the policy parameters, breaking the simple unbiasedness argument that holds in the synchronous setting. The hindsight-credit formulation in Eq.~\eqref{eq:hindsight_reward_app} is a coarser but markedly more stable alternative: it injects per-trip execution outcome (gap regularity) into upper credit assignment without requiring lower advantage backflow.

\subsection{What changes for our off-policy SAC upper}
The TransitDuet upper is trained as RE-SAC (\cref{sec:upper}, Eqs.~\ref{eq:upper_q_target}--\ref{eq:upper_policy_loss}), not as REINFORCE: each per-dispatch transition is pushed to a replay buffer, and the policy/critic are updated by repeated mini-batch SAC gradient steps. The REINFORCE-based unbiasedness argument that applies to a synchronous on-policy estimator is therefore not the right analytical lens for our setup. In practice the lower is also being updated between upper SAC gradient steps, so the upper's training distribution is non-stationary; the goal-conditioned coupling (\cref{sec:tap}) and TPC's importance reweighting (\cref{sec:tpc}) together control this non-stationarity at the lower side, and the upper's ensemble-LCB Q value provides conservatism against epistemic uncertainty in the upper's own off-policy training.
